---
title: "Exploring Datasets"
output: learnr::tutorial
tutorial:
  id: "Lab_catvars"
  version: 2.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Exploring_Datasets-T.html")){
  file.remove("Exploring_Datasets-T.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
  Survived = factor(Survived)
)

titanic_full <- titanic %>% mutate(
   Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
   Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)

### %ni = new input (red), %oi = old input (blue)
```


## Checking out the data

For this lab, the `titanic` dataset is already loaded for you.  The first thing we generally want to do is familiarize ourselves with the data. You can check out the first few rows of your dataset using the function `head()`.  You can also see how many rows the dataset has using `nrow()`.

When we want to do something to our dataset, we use a *pipe*, which looks like `%>%`.  You can think of this as "sending" the dataset into a process.

```{r, context="server"}

bob <- makePrintText(
          base_string = "%oi %>% head() \n %oi %>% nrow()",
          old_input = c("titanic", "titanic")
        )

output$code_1_2 <- renderText(bob)

output$code_1_3 <- renderText(
        makePrintText(
          base_string = "head(%oi) \n nrow(%oi)",
          old_input = c("titanic", "titanic")
        )
  )
```

```{r}
htmlOutput("code_1_2")
```


Note that we can also simply put the name `titanic` inside the parentheses of the function, like this:

```{r}
htmlOutput("code_1_3")
```

However, it is good to get into the habit of using the pipe - as we will see a bit later on, this will become convenient when we want to use many functions in a row on the same dataset.

Check out the result of the above lines of code:

```{r, echo = TRUE}
titanic %>% head()
titanic %>% nrow()
```

### Questions:

```{r q_1.1}
textInput("a_1.1.1",
           "How many Titanic passengers are represented in this dataset?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.1")

textInput("a_1.1.2",
           "How old was Miss Laina Heikkinen?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.2")

textInput("a_1.1.3",
           "How much did Mr. William Henry Allen pay for his ticket, in dollars?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.3")

radioButtons("a_1.1.4",
           "Did Mr. James Moran survive the sinking of the Titanic?",
           choices = c("Don't know" = "dk",
              "Yes" = "y",
             "No" = "n"))

htmlOutput("check_1.1.4")
```

```{r, context = "server"}

output$check_1.1.1 <- renderText(
  if(input$a_1.1.1 == ""){
      ""
  }else if(input$a_1.1.1 == nrow(titanic)){
     correct
  }else{
    incorrect
  }
)

output$check_1.1.2 <- renderText(
  if(input$a_1.1.2 == ""){
     ""
  }else if(input$a_1.1.2 == 26){
      correct
  }else{
    incorrect
  }
)

output$check_1.1.3 <- renderText(
  if(input$a_1.1.3 == ""){
     ""
  }else if(input$a_1.1.3 == 8.05){
      correct
  }else{
    incorrect
  }
)

output$check_1.1.4 <- renderText(
  if(input$a_1.1.4 == ""){
     ""
  }else if(input$a_1.1.4 == "n"){
      correct
  }else{
    incorrect
  }
)


```

## Looking at part of the dataset

Sometimes, we are only interested in looking at one piece of the dataset at a time.  

### Choosing columns

Perhaps we only want to consider a few of the variables. This can be done using `select()`.  Suppose we want to look at the name of the passenger, and only one other piece of information.  We will select these two variables, and then *pipe* that data to `head()` so we can look at the first few values.

```{r select}
textInput("select_var",
          "What other variable would you like to look at?",
          value = "",
          width = '80%', placeholder = NULL)

```

```{r, context="server"}
output$code_select <- renderText(
        makePrintText(
          base_string = "%oi %>% select(%oi, %ni) %>% head() ",
          old_input = c("titanic", "Name"),
          new_input = c(input$select_var)
        )
  )

output$output_select <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = "%oi %>% select(%oi, %ni) %>% head()",
          old_input = c("titanic", "Name"),
          new_input = c(input$select_var)
    )
   ))
})
```

Your code:

```{r}
htmlOutput("code_select")
```

Your output:
```{r}
verbatimTextOutput("output_select")
```


### Choosing rows

More often, we want to narrow down our sample according to one of the variables.  In **R**, we do this using `filter()`.  

Suppose we want to see only the passengers who survived the sinking.


```{r filter}
textInput("filter_var",
          "What variable contains the information about who survived the sinking?",
          value = "",
          width = '80%', placeholder = NULL)

radioButtons("filter_val",
             "What value do you want that variable to have?",
             choices = c("0" = "0",
               "1" = "1"))

```

```{r, context="server"}
output$code_filter <- renderText(
        makePrintText(
          base_string = "%oi %>% filter(%ni == %ni) %>% head()",
          old_input = c("titanic"),
          new_input = c(input$filter_var, input$filter_val)
        )
  )

output$output_filter <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = "%oi %>% filter(%ni == %ni) %>% head()",
          old_input = c("titanic"),
          new_input = c(input$filter_var, input$filter_val)
    )
   ))
})
```

Your code: (notice the double "equals" signs!)

```{r}
htmlOutput("code_filter")
```

Your output:
```{r}
verbatimTextOutput("output_filter")
```

```{r q_filter}
textInput("a_filter",
           "What is the first name of the 6th survivor listed?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_filter")
```

```{r, context = "server"}
output$check_filter <- renderText(
  if(input$a_filter == ""){
     ""
  }else if(input$a_filter == "Marguerite"){
      correct
  }else{
    incorrect
  }
)


```


## Making new variables

It is very rare that data comes to us in exactly the format we hope, or that **R** interprets the variable exactly as it should.  You will nearly always need to adjust the data a bit before you begin calculating or plotting anything.  

You, as the scientist, will need to decide how to create each new variable.  This will depend on your **research question** and your choice of **analysis**. 

### Adding two variables

Suppose we are interested in combining information in the dataset.  For example, our `titanic` data contains two pieces of information:  The number of siblings or spouses travelling with a passenger, and the number of parents or children travelling with a passenger.

We may not be interested in this distinction; instead, we want to know how many family members overall are travelling with each passenger.  Let's choose to name this variable `Family.Aboard`.

To create this, we will `mutate()` our dataset, which means adding or changing a variable. 


```{r add_vars}
textInput("sib_spouse",
          "What is the name of the variable containing the number of siblings or spouses travelling with each passenger?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("parent_child",
          "What is the name of the variable containing the number of parents or children travelling with each passenger?",
          value = "",
          width = '80%', placeholder = NULL)


```

```{r, context="server"}

output$code_fam <- renderText(
        makePrintText(
          base_string = "%oi <- %oi %>% mutate( \n \t %oi = %ni + %ni \n )",
          old_input = c("titanic", "titanic", "Family.Aboard"),
          new_input = c(input$sib_spouse, input$parent_child)
        )
  )


```

```{r}
titanic <- titanic %>% mutate(
  Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard
)
```

The code for this will look like:

```{r}
htmlOutput("code_fam")
```

Notice how we always use `<-` to update the dataset, i.e., to save the new calculations into our data.  Then, if we look at the data again, the new variable should be there:

```{r}
titanic %>% head()

```


### Quantitative and Categorical variables

Consider the variable called `Pclass`.  As you hopefully realized, this variable should be **categorical** - it tells which class (First, Second, or Third) the passenger travelled in.  However, here is what **R** thinks about the *class*, or type, of that variable:

```{r, echo = TRUE}
class(titanic$Pclass)
```

It is an "integer" - that is, a discrete, quantitative variable.  Oh no!

Let's make a new variable named `Passenger.Class`, using `mutate()`.  The function `factor()` will turn a numeric variable into categories.  


```{r factor_var}
textInput("pclass",
          "What is the name of the existing variable that you would like to make categorical?",
          value = "",
          width = '80%', placeholder = NULL)


```

```{r, context="server"}

output$code_3_1 <- renderText(
        makePrintText(
          base_string = "%oi <- %oi %>% mutate( \n \t %oi = factor(%ni) \n )",
          old_input = c("titanic", "titanic", "Passenger.Class"),
          new_input = c(input$pclass)
        )
  )



```

The code for this will look like:

```{r}
htmlOutput("code_3_1")
titanic <- titanic %>% mutate(
  Passenger.Class = factor(Pclass)
)
```

And we can check to make sure our new variable exists and did what we hoped:

```{r, echo = TRUE}
class(titanic$Passenger.Class)

```


Perhaps you think it might be convenient to have the values of your new variable be more informative than `1` and `2` and `3`.  This is easy to do with `factor()`.

```{r factor_var_names}
textInput("word_1",
          "What label would you like to use for 1?",
          value = "",
          width = '80%', placeholder = NULL)


textInput("word_2",
          "What label would you like to use for 2?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("word_3",
          "What label would you like to use for 3?",
          value = "",
          width = '80%', placeholder = NULL)


```

```{r, context="server"}

output$code_3_2 <- renderText(  
  makePrintText(
          base_string = "titanic <- titanic %>% mutate( \n \t %oi = factor(%oi, \n \t \t levels = c(1,2,3), \n \t \t labels = c('%ni', '%ni', '%ni')) \n )",
          old_input = c("Passenger.Class", input$pclass),
          new_input = c(input$word_1, input$word_2, input$word_3)
        ))
```

```{r}
htmlOutput("code_3_2")
titanic <- titanic %>% mutate( 
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third")))
```

And check your dataset:

```{r, echo = TRUE}

titanic %>% select(Name, Pclass, Passenger.Class) %>% head()

```

This is your first experience with optional arguments in a function!  Notice that `factor()` worked just fine with only a variable name, but that you were able to do even more with the options `levels` and `labels`.