---
title: "Linear Regression"
output: learnr::tutorial
tutorial:
  id: "Lab_regression"
  version: 2.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_Linear_Regression.html")){
  file.remove("Lab_Exercise_Linear_Regression.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)
require(tibble)
require(fmsb)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

babies = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/babies_sub.csv')
kellys <- babies %>% 
  filter(Name == "Kelly") %>%
  mutate(
    Name = factor(Name)
  )

### %ni = new input (red), %oi = old input (blue)
```

## Baby Names in the U.S.

Our dataset in this lab studies baby names in California.  The loaded dataset, `kellys`, has information about how many babies were given the name "Kelly" in California from 1940-2016.  

Take a look at the dataset, and familiarize yourself with the variables.

```{r ex_1-1, exercise = TRUE}
```


## Scatterplots

Let's take a look at how the name "Kelly" has changed over time, for female-assigned babies only.

Use the following code to plot the counts of babies named Kelly (in California) for each year.  Notice the new addition to `ggplot()` of `+ geom_point()`


```{r, echo=FALSE}
textInput("plot_var_2_1", 
          "What variable would you like to plot on the x-axis?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_2_2", 
          "What variable would you like to plot on the y-axis?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

radioButtons("plot_type", 
             "Would you like a scatterplot or a line?",
                   choices = c("Scatterplot" = "point",
                               "Line" = "line"
                              )
              )

# 
htmlOutput("code_2")
# 
plotOutput("plot_2")
```


```{r, context="server"}
bs_2 <- reactive("%oi %>% \n filter(Gender == 'F') %>% \n ggplot(aes(x = %ni, y = %ni)) + geom_%ni()")

output$plot_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_2(), 
      old_input = c("kellys"),
      new_input = c(input$plot_var_2_1, input$plot_var_2_2, input$plot_type)
    )
   ))
})

output$code_2 <- renderText(
        makePrintText( 
      base_string = bs_2(), 
      old_input = c("kellys"),
      new_input = c(input$plot_var_2_1, input$plot_var_2_2, input$plot_type)
    )
  )

```

**Think about it:**  Why did the name "Kelly" suddenly become popular for female babies around 1960?  I have heard a few theories... can you find me an interesting one?  

## Regression

It appears that the name "Kelly" is getting steadily less popular since 1980.  However, we would like to know if this trend is significant.  First, let's adjust our dataset to narrow it down to only female babies and only 1980 and beyond.

```{r, eval = TRUE, echo = TRUE}
kellys_80 <- kellys %>% filter(Gender == "F", Year > 1980)
```

Use the code below to run a **linear regression** on the counts of Kellys over time.  For this, we use the **R** command `lm()`, which stands for "linear model". 



```{r, echo=FALSE}
textInput("plot_var_3_1", 
          "What is the explanatory variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_3_2", 
          "What is the response variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


# 
htmlOutput("code_3")
# 
verbatimTextOutput("plot_3")
```


```{r, context="server"}
bs_3 <- reactive("summary(%oi(%ni ~ %ni, data = %oi))")

kellys_80 <- kellys %>% filter(Gender == "F", Year > 1980)

output$plot_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c("lm", "kellys_80"),
      new_input = c(input$plot_var_3_2, input$plot_var_3_1)
    )
   ))
})

output$code_3 <- renderText(
        makePrintText( 
      base_string = bs_3(), 
      old_input = c("lm", "kellys_80"),
      new_input = c(input$plot_var_3_2, input$plot_var_3_1)
    )
  )

```


```{r q_3}
textInput("a_3",
           "What is the test statistic (F) for this regression line?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_3")

```

```{r, context = "server"}

output$check_3 <- renderText(
  if(round(as.numeric(input$a_3), 2) == 1565){
      correct
  }else{
    incorrect
  }
)
```


```{r q_31}
textInput("a_31",
           "What is the test statistic (t) for the slope of this regression line?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_31")

```

```{r, context = "server"}

output$check_31 <- renderText(
  if(round(as.numeric(input$a_31), 2) == -39.55){
      correct
  }else{
    incorrect
  }
)
```

```{r q_4}
textInput("a_4",
           "What was the p-value for this regression line?",
           value = "1",
           width = '80%', placeholder = NULL)

htmlOutput("check_4")

```

```{r, context = "server"}

output$check_4 <- renderText(
  if(as.numeric(input$a_4) == 0){
      correct
  }else{
    incorrect
  }
)
```



```{r q_6}
textInput("a_6",
           "What percent of variance in the number of Kellys born in a year is explained by a change over time (r-squared)?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_6")

```

```{r, context = "server"}

output$check_6 <- renderText(
  if(round(as.numeric(input$a_6), 2) == 97.87){
      correct
  }else{
    incorrect
  }
)
```


```{r q_5}
textInput("a_5",
           "What is the correlation between number of Kellys and time (r)?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_5")

```

```{r, context = "server"}

output$check_5 <- renderText(
  if(round(as.numeric(input$a_5), 2) == -0.99){
      correct
  }else{
    incorrect
  }
)
```

## Checking residuals

The last thing we need to do is make sure the *linear* regression is a good choice, rather than some other shape of line.  Use the code below to plot the regression line.

```{r, echo=FALSE}
textInput("plot_var_4_1", 
          "What is your explanatory variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_4_2", 
          "What is your response variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


# 
htmlOutput("code_4")
```


```{r, context="server"}
bs_4 <- reactive("my_lm <- lm(%ni ~ %ni, data = %oi)")

output$code_4 <- renderText(
        makePrintText( 
      base_string = bs_4(), 
      old_input = c("kellys_80", "point"),
      new_input = c(input$plot_var_4_1, input$plot_var_4_2, input$reg_line)
    )
  )

```

The "residual" of an observation is how far away the prediction on the regression line is from what we actually saw.  We hope that there is no clear pattern in the residuals; i.e., that any "wrong" prediction is due purely to chance.

Here is a plot of the residuals.


```{r, echo = FALSE}
my_lm <- lm(Count ~ Year, data = kellys_80)
```

```{r, echo = TRUE}

ggplot(my_lm) + 
  geom_point(aes(x=.fitted, y=.resid)) +
  xlab("Fitted values") + ylab("Residuals") +
  geom_hline(yintercept = 0, col = "red")

```

```{r, echo=FALSE}
textAreaInput("discuss_1", 
          "Do you see any issue with the regression, based on the residuals?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

```
