---
title: "Lab 4: Video Games"
author: "Stat 218"
output: rmdformats::readthedown
---
## Setup

```{r, include = FALSE}
library(tidyverse)
```


Create a fresh new R Markdown file.  Add in your `library()` statements, if you are using the tidyverse.  Read in your data:

```{r}
vgames <- read_csv("https://www.dropbox.com/s/yaqtky7f6v89unq/video_games.csv?dl=1")
```



## The Dataset

This analysis is an exploration of video games published over the past 20(ish) years.

Notes:

* The "Platform" of a video game is the machine that it is designed to be played on.
* The "Publisher" is the company that created the game
* "NA_Sales" stands for North America, "EU" is European Union, and "JP" is Japan.
* Sales are measured in millions of copies sold

Explore the dataset, and describe it briefly using class vocabulary.

## One Categorical Variable: Genre

A group of parents is concerned with the the rise in violent video games.  This group claims that  10% of all video games published in 2020 will be in the genre of "Shooter".  Does the data support this claim?

* **Research Question:**

* **Parameter(s):** 
    + If yes:
    + If no:
    
* **Summary Statistic Sentence:**

* **Plot:**

* **Test Statistic:**

* **Conclusion:**

```{r}
vgames %>%
  count(Genre)

vgames %>%
  nrow()

1095/14003
```




## One Quantitative Variable:  Nintendo Sales

*Nintendo* is a major video game publisher.  One expert suggests that Nintendo might average over 3 million sales per game over the next 20 years.  Does the data support this prediction?

## Two Categorical Variables: Genre and Publisher

Consider the game publishers "Nintendo" and "Electronic Arts (EA)".  Many people consider EA to be a company known for its sports games. Does the data support this characterization?

Hint: Reduce your datset to only Nintendo and EA.  Then create a new variable called `is_sports` that has variables `TRUE` or `FALSE`.  Then count up your variables.  Here is some code to help you out:

```{r}
vgames %>%
  filter(Publisher %in% c("Nintendo", "Electronic Arts")) %>%
  mutate(
    is_sports = (Genre == "Sports")
  )
```


## One Categorical and One Quantitative:  Sales in Europe and North America

Does North America tend to buy more video games than Europe?  Make a case based on the data.

## Bonus:  Two Quantitative Variables:  Games over time

**THIS SECTION IS EXTRA CREDIT**

Have the Global sales of video games increased over time?  Make visualization, and use it to argue for your answer.