---
title: "Lab 5: Cal Poly Basketball revisited"
author: "Stat 218"
output: rmdformats::readthedown
---
## Setup

```{r, include = FALSE}
library(tidyverse)
```

In your **console** (bottom left), NOT your Markdown document, type

```{r, eval = FALSE}
install.packages("infer")
```


Then create a fresh new R Markdown file.  Add in your `library()` statements, if you are using the tidyverse.  This time, you should also add `library(infer)`.

In this lab we will revisit the Cal Poly Basketball dataset:

```{r, message = FALSE}
bball = read.table("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/cp_bball.csv", header = TRUE)
```

We will now answer our questions from Lab 3 using hypothesis testing.

## Is Cal Poly a worse than average team?

An "average" team is defined to be a team that is just as likely to lose as to win a game.  We would like to determine if Cal Poly's team was average or not.  

In Lab 3, we found the following data:

```{r}
bball %>%
  count(Team.Result)
```

Fill in the blanks in the following (you do not need to show work in code):

Under the **null hypothesis**, we expect the **true percentage** of wins to be [_____].

Under the **alternate hypothesis**, we expect the **true percentage** of wins to be [_____].

The **sample proportion** in this data is [_____].

The **standard deviation** of the sample proportion is [______].

How many standard deviations was what we saw from what we expected?  Save this as "t".

What is the sample size?  Save this as "n"

```{r, eval = FALSE}
t <- ### number of sds
n <- ## sample size
```

Then run the code below without changing it:

```{r, eval = FALSE}
pt(t, df = n)
```

This is your p-value!

What do you conclude?


## Was Cal Poly more likely to win at home than away?

Perhaps Cal Poly performs better in front of a home crowd.  Let's find out.

Perform a permutation test, like you did in the Lab Exercise, to make a conclusion.


## Does Cal Poly score more at home than away?

Perhaps Cal Poly scores more in front of a home crowd, regardless of wins and losses.


### Hypothesis test using standard deviation

Write down your null and alternate hypotheses.

The observed difference of mean scores is [------]

The standard error of the observed difference of mean scores is [------].

The number of standard errors between what we saw and what we expected was [-----].  Save this number as "t"

The sample size in the larger category (either "home" or "away") was [----].  Save this number as "n".

```{r, eval = FALSE}
t <- ### number of sds
n <- ## sample size
```

Then run the code below without changing it:

```{r, eval = FALSE}
pt(t, df = n)
```

What do you conclude?

### Hypothesis testing using permutation

Perform a permutation test, like you did in the Lab Exercise.  How does your p-value compare to the one above?
