---
title: "Lab 7: Baby Names"
author: "Stat 218"
output: rmdformats::readthedown
---
## Partners

You **may** work in teams of two for this lab, if you wish.  Simply turn in the same .html file for both members of the team.

**Be sure to put both author's names on the document, so I know you worked together.**

## Setup

### 1.  

Choose a name (or several) that you wish to study.

### 2.  

Navigate to the data:

https://www.dropbox.com/sh/czm61yzmvzl6xfb/AABEFdpy02aNRhZ_vcxrVAQSa?dl=0 .

### 3.  

Download to you computer the dataset(s) you need for the name(s) you are studying.  **Make sure you save these datasets in your Stat 218 folder!!!**

### 4.  

Create a new R Markdown file; add up your `library(tidyverse)` statement.

### 5.  

Read your dataset(s) into your file as follows:

```{r, eval = FALSE}
k_names <- read.csv("StateNames_K.csv")
```

If this step does not work:

* Make sure your data file (the `.csv` one) is in your Stat 218 folder.

* Make sure the name of the data file is "StateNames_K.csv"

* Double check the letter - in my example, it is `K`, but you may have picked different letters!


### 6. 

If you are studying many names, and you wish to combine the datasets into one big dataset, use the following code (possibly changing `k_names` and `b_names` to match your own datasets:

```{r, eval = FALSE}
all_names <-
  k_names %>% bind_rows(b_names)
```

##  Research Question

Use either an ANOVA analysis or a Linear Regression analysis to reach a conclusion about your chosen name(s), as in the Lab Exercises.
