---
title: "Visualizing Categorical Variables"
output: learnr::tutorial
tutorial:
  id: "Lab_catvars"
  version: 2.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Summarizing_Categorical_Variables-T.html")){
  file.remove("Summarizing_Categorical_Variables-T.html")
}


require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
  Survived = factor(Survived)
)

titanic <- titanic %>% mutate(
   Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
   Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)

### %ni = new input (red), %oi = old input (blue)
```

## Introduction

In this exercise, we will begin visualizing the information from the `titanic` data.  We will start by exploring the **categorical** variables in the dataset.

First, let's make the necessary adjustments to the categorical variables, so that **R** recognizes them as categorical (or "factors").

```{r, echo = TRUE}
titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
```

## Intro to ggplot

We will be using a package called `ggplot2` to make all of our visualizations.  This package has a very specific format (or "grammar of graphics") for creating plots:  

### Mapping

We first must specify the **mapping** or **aesthetic**.  This is where we "map" certain variables to the place on the chart where they will be visualized.

### Geometry

We then must specify the **geometry**, or the type of plot we would like to make.  We can also add various options to make the plot prettier.


Read on to try this out:


## Bar Plots

```{r, echo=FALSE}
textInput("plot_var_1", 
          "What variable would you like to plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_dat_1", 
          "What dataset is that variable found in?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

# radioButtons("plot_type_1",
#              "Would you like a barplot or a dotplot?",
#              choices = c("barplot" = "bar",
#                          "dotplot" = "dotplot"))


htmlOutput("code_4_1")

plotOutput("plot_4_1")
```


```{r, context="server"}
bs_4_1 = reactive("ggplot(%ni, aes(x = %ni)) + geom_bar()")

output$plot_4_1 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_1(), 
      new_input = c(input$plot_dat_1, input$plot_var_1)
    )
   ))
})

output$code_4_1 <- renderText(
        makePrintText( 
          base_string = bs_4_1(), 
          new_input = c(input$plot_dat_1, input$plot_var_1)
        )
  )

```

Beyond the crucial information of the dataset and the variable(s) to plot, there are also plenty of optional settings to make our image nicer.

```{r, echo=FALSE}
textInput("plot_title_1", 
          "Title of the plot:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_xlab_1", 
          "Label for x-axis:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

radioButtons("fill_1", 
             "Should the bars be different colors?",
                   choices = c("No" = "",
                               "Yes" = ", fill = %oi")
              )


```

```{r, context = "server"}
bs_4_2 <- reactive("ggplot(%oi, aes(x = %oi%ni)) + geom_bar() + ggtitle('%ni') + xlab('%ni')")

output$plot_4_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_2(), 
      old_input = c("titanic", input$plot_var_1, input$plot_var_1),
      new_input = c(input$fill_1, input$plot_title_1, input$plot_xlab_1)
    )
   ))
})

output$code_4_2 <- renderText(
        makePrintText( 
          base_string = bs_4_2(), 
          old_input = c(input$plot_dat_1, input$plot_var_1, input$plot_var_1),
          new_input = c(input$fill_1, input$plot_title_1, input$plot_xlab_1)
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_2")

plotOutput("plot_4_2")
```

## Multiple bar plots

Now suppose we wish to compare two categorical variables.  One option is to simply make separate barplots for each category.  We do so by adding the option `facet_grid()` to our plot.  Try it below.

```{r, echo=FALSE}

textInput("plot_var_2_1", 
          "What variable would you like use for the bars?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_2", 
          "What variable would you like to use to separate the plots?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_title_2", 
          "Title of the plot:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_xlab_2", 
          "Label for x-axis:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

radioButtons("fill_2", 
             "Should the bars be different colors?",
                   choices = c("No" = "",
                               "Yes" = ", fill = %oi")
              )





```

```{r, context = "server"}
bs_4_3 <- reactive("ggplot(%oi, aes(x = %ni%ni)) + geom_bar() + facet_grid(~%ni) + ggtitle('%ni') + xlab('%ni')")

output$plot_4_3 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_3(), 
      old_input = c("titanic", input$plot_var_2_1),
      new_input = c(input$plot_var_2_1, input$fill_2, input$plot_var_2, input$plot_title_2, input$plot_xlab_2)
    )
   ))
})

output$code_4_3 <- renderText(
        makePrintText( 
      base_string = bs_4_3(), 
      old_input = c("titanic", input$plot_var_2_1),
      new_input = c(input$plot_var_2_1, input$fill_2, input$plot_var_2, input$plot_title_2, input$plot_xlab_2)
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_3")

plotOutput("plot_4_3")
```

## Side-by-side bar plots

Another option is to plot bars side-by-side, separating groups by the value of the second categorical variable.
```{r, echo=FALSE}

textInput("plot_var_3_1", 
          "What variable would you like use for the bar heights?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_3_2", 
          "What variable would you like to use to separate the bars?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_title_3", 
          "Title of the plot:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_xlab_3", 
          "Label for x-axis:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)





```

```{r, context = "server"}

bs_4_4 <- reactive("ggplot(%oi, aes(x = %ni, fill = %ni)) + geom_bar(%ni) + ggtitle('%ni') + xlab('%ni')")

output$plot_4_4 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_4(), 
      old_input = c("titanic"),
      new_input = c(input$plot_var_3_1, input$plot_var_3_2, "position = 'dodge'", input$plot_title_3, input$plot_xlab_3)
    )
   ))
})

output$code_4_4 <- renderText(
        makePrintText( 
          base_string = bs_4_4(), 
      old_input = c("titanic"),
      new_input = c(input$plot_var_3_1, input$plot_var_3_2, "position = 'dodge'", input$plot_title_3, input$plot_xlab_3)
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_4")

plotOutput("plot_4_4")
```

## Stacked Bar Charts

The last, and typically best option is to make a **stacked bar chart**, which combines the two categorical variables on the bars themselves.  Note that these charts continue to obey the *area principle* - the area of the colored sections corresponds to the number of people in that category!

Another advantage of the stacked bar chart is that it is a great alternative to a pie chart for looking at percentages.  Try plotting the bar heights as percentages rather than counts.  Think about what research questions might be better addressed by this approach.

```{r, echo=FALSE}

textInput("plot_var_4_1", 
          "What variable would you like use for the bar heights?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_4_2", 
          "What variable would you to color the bar stacks?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_title_4", 
          "Title of the plot:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_xlab_4", 
          "Label for x-axis:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

radioButtons("plot_pct",
             "The height of the bars should be:",
             choices = c(
               "Counts" = "",
               "Percentages" = "position = 'fill'"
             ))






```

```{r, context = "server"}
bs_4_5 <- reactive("ggplot(%oi, aes(x = %ni, fill = %ni)) + geom_bar(%ni) + ggtitle('%ni') + xlab('%ni')")

output$plot_4_5 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_5(), 
      old_input = c("titanic"),
      new_input = c(input$plot_var_4_1, input$plot_var_4_2, input$plot_pct, input$plot_title_4, input$plot_xlab_4)
    )
   ))
})

output$code_4_5 <- renderText(
        makePrintText( 
          base_string = bs_4_5(), 
      old_input = c("titanic"),
      new_input = c(input$plot_var_4_1, input$plot_var_4_2, input$plot_pct, input$plot_title_4, input$plot_xlab_4)
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_5")

plotOutput("plot_4_5")
```


Play around with these various plots, until you find an insight that you think is interesting!
