---
title: "Summary Statistics"
output: learnr::tutorial
tutorial:
  id: "Lab_Summaries"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Summary_Statistics-T.html")){
  file.remove("Summary_Statistics-T.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)

### %ni = new input (red), %oi = old input (blue)
```


## Introduction

In this exercise, we will continue to explore the `titanic` dataset.  The dataset has already been loaded, cleaned, and adjusted.

Instead of exploring the data and answering research questions with *visualizations*, we will now learn to calculate *summary statistics*.

Remind yourself what the dataset looks like:

```{r, echo = TRUE}
titanic %>% head()
```

## General Summaries

An easy method, and a good first step in any data analysis, is to use the function `summary()` to get some basic information about all the variables:


```{r, echo = TRUE}
titanic %>% summary()
```

## Categorical Variables

To summarize a categorical variable, we would like to know how many observations occurred in each category.  This is easily achieved with `count()`.


```{r count}
textInput("var_00",
          "What is the categorical variable you would you like to count?",
          value = "",
          width = '80%', placeholder = NULL)




```

```{r, context="server"}

bs_00 = reactive("%oi %>% count(%ni)")

output$code_00 <- renderText(
        makePrintText(
          base_string = bs_00(),
          old_input = c("titanic"),
          new_input = c(input$var_00)
        )
  )

output$output_00 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_00(),
          old_input = c("titanic"),
          new_input = c(input$var_00)
    )
   ))
})

```

```{r}
htmlOutput("code_00")
```

Result:

```{r}
verbatimTextOutput("output_00")
```


### More than one categorical variable

If we would like to find counts in multiple categories, we can simply add arguments:

```{r count2}
textInput("var_0_1",
          "What is the first categorical variable you would you like to count?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("var_0_2",
          "What is the second categorical variable you would you like to count?",
          value = "",
          width = '80%', placeholder = NULL)




```

```{r, context="server"}
bs_0 = reactive("%oi %>% count(%ni, %ni)")

output$code_0 <- renderText(
        makePrintText(
          base_string = bs_0(),
          old_input = c("titanic"),
          new_input = c(input$var_0_1, input$var_0_2)
        )
  )

output$output_0 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_0(),
          old_input = c("titanic"),
          new_input = c(input$var_0_1, input$var_0_2)
    )
   ))
})

```

```{r}
htmlOutput("code_0")
```

Result:

```{r}
verbatimTextOutput("output_0")
```

## Quantitative variables

Sometimes we are only interested in one specific summary statistic, or if we want to calculate something that is not included in `summary()`, such as **standard deviation**.  For this, we can use the function `summarize_at`.



```{r summarize}
textInput("var1",
          "Which QUANTITATIVE variable would you like to find information about?",
          value = "",
          width = '80%', placeholder = NULL)

checkboxGroupInput("funs1", 
          "What information do you want?",
          choices = c("Mean" = "mean",
                      "Median" = "median",
                      "Standard Deviation" = "sd",
                      "IQR" = "IQR",
                      "Sample Size" = "length")
)




```

```{r, context="server"}
bs_1_1 = reactive("%oi %>% summarize_at(vars(%ni), funs(%ni))")

output$code_1_1 <- renderText(
        makePrintText(
          base_string = bs_1_1(),
          old_input = c("titanic"),
          new_input = c(input$var1, toString(input$funs1))
        )
  )

output$output_1_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_1(), 
          old_input = c("titanic"),
          new_input = c(input$var1, toString(input$funs1))
    )
   ))
})

```

```{r}
htmlOutput("code_1_1")
```

Result:

```{r}
verbatimTextOutput("output_1_1")
```

## Quantitative variable by group

Some research questions will involve understanding how a quantitative variable relates to a categorical variable.  We may wish to calculate summary statistics separately for each category. We can achieve this with `group_by()`

```{r grooupby}
textInput("var2_quant",
          "Which QUANTITATIVE variable would you like to find information about?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("var2_cat",
          "Which CATEGORICAL variable would you like to use to split groups?",
          value = "",
          width = '80%', placeholder = NULL)

checkboxGroupInput("funs2", 
          "What information do you want?",
          choices = c("Mean" = "mean",
                      "Median" = "median",
                      "Standard Deviation" = "sd",
                      "IQR" = "IQR",
                      "Sample Size" = "length")
)




```

```{r, context="server"}

bs_1_2 = reactive("%oi %>% group_by(%ni) %>% summarize_at(vars(%ni), funs(%ni))")

output$code_1_2 <- renderText(
        makePrintText(
          base_string = bs_1_2(),
          old_input = c("titanic"),
          new_input = c(input$var2_cat, input$var2_quant, toString(input$funs2))
        )
  )

output$output_1_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
        base_string = bs_1_2(),
        old_input = c("titanic"),
        new_input = c(input$var2_cat, input$var2_quant, toString(input$funs2))
    )
   ))
})

```

```{r}
htmlOutput("code_1_2")
```

Result:

```{r}
verbatimTextOutput("output_1_2")
```



### Check your work

*(round to the nearest one decimal)*

```{r q_1.1}
textInput("a_1.1.1",
           "What is the standard deviation of the age of First Class passengers?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.1")

textInput("a_1.1.2",
           "What was the median age of Female passengers?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.2")

textInput("a_1.1.3",
           "What was the IQR for a ticket on the Titanic, based on this dataset?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.3")
```

```{r, context = "server"}

output$check_1.1.1 <- renderText(
  if(input$a_1.1.1 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1.1), 2) == round(sd(titanic[titanic$Pclass == 1, "Age"]),1)){
      correct
  }else{
    incorrect
  }
)

output$check_1.1.2 <- renderText(
   if(input$a_1.1.2 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1.2),2) == round(median(titanic[titanic$Sex == "female", "Age"]),1)){
      correct
  }else{
    incorrect
  }
)

output$check_1.1.3 <- renderText(
   if(input$a_1.1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1.3),2) == round(IQR(titanic$Fare, na.rm = TRUE), 1)){
      correct
  }else{
    incorrect
  }
)


```
