---
title: "Visualizing Quantitative Variables"
output: learnr::tutorial
tutorial:
  id: "Lab_QuantVar"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Visualizing_Quantitative_Variables-T.html")){
  file.remove("Visualizing_Quantitative_Variables-T.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)

### %ni = new input (red), %oi = old input (blue)
```


## Introduction

In this exercise, we will continue to explore the `titanic` dataset, which has already been loaded for you.  This time, we will think about **quantitative** variables - how to display them and how to answer research questions with plots.

First, we will make the necessary adjustments to the categorical variables:

```{r, echo = TRUE, eval = FALSE}
titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
```

Then, remind yourself what the dataset looks like:

```{r, echo = TRUE}
titanic %>% head()
```


## Histograms

To make other plots beyond the **barplots** that you already learned, you will use almost the same code.  We still use `ggplot()` in the same way, as well as options like `ggtitle()` and `xlab`.  The only difference is that instead of `geom_bar()`, we want to make a **histogram**, or `geom_histogram()`.  Try it below.  Pay careful attention to the new options that come with `geom_histogram()`

(Hint: Go to [This Page](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) to see lots of fun color options!)

```{r, echo=FALSE}
textInput("plot_2_var", 
          "What QUANTITATIVE variable would you like to make a histogram of?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

sliderInput("plot_2_bins", 
          "How many bins should the histogram have?",
          value = 20,
          min = 1, 
          max = 100)

textInput("plot_2_color", 
          "What color should the histogram be?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_2_title", 
          "What title will you give your plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_2_xlab", 
          "What label should the x-axis have?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)




# 
htmlOutput("code_2")
# 
plotOutput("plot_2")
```


```{r, context="server"}
bs_2 <- reactive("ggplot(%oi, aes(x = %ni)) + geom_%oi(bins = %ni, %ni) + ggtitle('%ni') + xlab('%ni')")

output$plot_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_2(), 
      old_input = c("titanic", "histogram"),
      new_input = c(input$plot_2_var, input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab)
    )
   ))
})

output$code_2 <- renderText(
        makePrintText( 
      base_string = bs_2(), 
      old_input = c("titanic", "histogram"),
      new_input = c(input$plot_2_var, input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab)
    )
  )

```

### Densities

A histogram is our best guess about the shape of a **density**.  We can also choose to plot a smooth line over the histogram, to help us visualize what the density might be like.

```{r}

radioButtons("count_density",
             "The height of the bars should be:",
             choices = c("counts" = "", 
                         "densities" = "aes(y = ..density..), ")
)

radioButtons("add_density",
             "The plot should show an estimated density curve.",
             choices = c(
                "no" = "",
               "yes" = "+ geom_density(size = 2)"
             )
)
               



# 
htmlOutput("code_3")
# 
plotOutput("plot_3")
```

```{r, context="server"}

bs_3 <- reactive("ggplot(%oi, aes(x = %oi)) + geom_%oi(%nibins = %oi%oi) %ni+ ggtitle('%oi') + xlab('%oi')")

output$plot_3 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c("titanic", input$plot_2_var, "histogram", input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab),
      new_input = c(input$count_density, input$add_density)
    )
   ))
})

output$code_3 <- renderText(
        makePrintText( 
      base_string = bs_3(), 
      old_input = c("titanic", input$plot_2_var, "histogram", input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab),
      new_input = c(input$count_density, input$add_density)
    )
  )

```

### Reading the histogram

Use the exercise above to with the variable `Age`. Think about the following questions:

* What number of bins one seems to give the most reasonable density estimate, in your opinion?

* How many modes does this distribution have?

* Is this distribution skewed?

* What are the **mean** and **median** age of Titanic passengers?


## Multiple Histograms

As we did with barplots, we will use the option `facet_grid` to make multiple separate *histograms* across values a particular categorical variable.


```{r, echo=FALSE}
textInput("plot_4_var_quant", 
          "What QUANTITATIVE variable would you like to plot on a histogram?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_4_var_cat", 
          "What CATEGORICAL variable would you like to use to separate the histograms?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


# 
htmlOutput("code_4")
# 
plotOutput("plot_4")
```


```{r, context="server"}
bs_4 <- reactive("ggplot(%oi) + geom_histogram(aes(x = %ni)) + %oi(~%ni)")

output$plot_4 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4(), 
      old_input = c("titanic", "facet_grid"),
      new_input = c(input$plot_4_var_quant, input$plot_4_var_cat)
    )
   ))
})

output$code_4 <- renderText(
        makePrintText( 
      base_string = bs_4(), 
      old_input = c("titanic", "facet_grid"),
      new_input = c(input$plot_4_var_quant, input$plot_4_var_cat)
    )
  )

```

## Side-by-Side Boxplots

A better way to compare a quantitative and a categorical variable is usually to make **side-by-side boxplots**.


```{r, echo=FALSE}
textInput("plot_5_var_quant", 
          "What QUANTITATIVE variable would you like to plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_5_var_cat", 
          "What CATEGORICAL variable would you like to compare?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

# 
htmlOutput("code_5")
# 
plotOutput("plot_5")
```


```{r, context="server"}
bs_5 <- reactive("ggplot(%oi) + geom_%oi(aes(x = %ni, y = %ni, fill = %ni))")

output$plot_5 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_5(), 
      old_input = c("titanic", "boxplot"),
      new_input = c(input$plot_5_var_cat, input$plot_5_var_quant, input$plot_5_var_cat)
    )
   ))
})

output$code_5 <- renderText(
        makePrintText( 
      base_string = bs_5(), 
      old_input = c("titanic", "boxplot"),
      new_input = c(input$plot_5_var_cat, input$plot_5_var_quant, input$plot_5_var_cat)
    )
  )

```
